---
title: "December Psi Chi R Submission"
output: html_document
author: Brady M. Chisholm
---
### Prep & Setup
```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))

# cache
install.packages("knitr", quiet = TRUE)
library(knitr)
required_packages <- c("ggplot2","tidyr","readr","moments","psych","gt") 
# code into a function 
for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE)) { 
    install.packages(pkg)               
    library(pkg, character.only = TRUE)} 
  else {
    library(pkg, character.only = TRUE)}}
```
Load data 
```{r data loading, echo=TRUE, message=FALSE, warning=FALSE}
# Load CSV file
dataSource <- "https://raw.githubusercontent.com/bchisholm08/OSF_graphingCompetition/main/Month/shopping_trends_updated.csv"
data <- read_csv(dataSource)

# preview data / var types
head(data)
```

### Level 1 
```{r lvl 1 filtering}
# filter NA's in age
naRMData <- subset(data, !is.na(Age))

# Check filtered data struct 
head(naRMData)
```

```{r}
ggplot(naRMData, aes(x = `Review Rating`)) +
  geom_histogram(binwidth = 0.2, fill = "#dcfbe7", color = "black") +  
  labs(
    title = "Distribution of Review Rating Variable",
    x = "Review Rating",
    y = "Frequency"
  ) +
  theme_minimal(base_size = 15) + 
  theme(
    panel.background = element_rect(fill = "black", colour = "black"),  
    plot.background = element_rect(fill = "black", colour = "black"),   
    panel.grid.major = element_line(colour = "#fbe7dc"),                
    panel.grid.minor = element_line(colour = "#fbe7dc"),                
    axis.text = element_text(colour = "#dcfbe7"),                       
    axis.title = element_text(colour = "#e7dcfb", size = 14),           
    plot.title = element_text(colour = "#fbe7dc", size = 20, hjust = 0.5),  
    plot.margin = margin(10, 10, 10, 10))
```

Distribution of rating. Skew val 
```{r lvl 1 skew}
# Calculate skew
skewness_value <- moments::skewness(naRMData$`Review Rating`)
paste("Review Rating Skew =", round(skewness_value, 4))
```
>**Review rating data is normally distributed, and does not skew to the left or right**  
>**Skew value is low, only 0.0058**

### Level 2 
min, med, avg, sd of review rating
```{r summary stats}
descriptiveStats <- data.frame(
  Metric = c("Minimum", "Median", "Maximum", "Average", "Standard Deviation"),
  Value = c(
    min(naRMData$`Review Rating`, na.rm = TRUE),
    median(naRMData$`Review Rating`, na.rm = TRUE),
    max(naRMData$`Review Rating`, na.rm = TRUE),
    mean(naRMData$`Review Rating`, na.rm = TRUE),
    sd(naRMData$`Review Rating`, na.rm = TRUE)))

descriptiveStats %>%
  gt() %>%
  tab_header(
    title = "Descriptive Statistics for Review Rating"
  ) %>%
  fmt_number(
    columns = "Value",
    decimals = 3
  ) %>%
  cols_label(
    Metric = md("**Statistic**"),
    Value = md("**Value**"))
```

### Level 3 
table of avg rating by category
```{r lvl 3 tbl}
# pull table vals 
catAvg <- naRMData %>%
  dplyr::group_by(Category) %>%
  dplyr::summarise(Average_Review_Rating = mean(`Review Rating`, na.rm = TRUE)) %>%
  dplyr::arrange(desc(Average_Review_Rating))

# explore gt package...never tried it
catAvg %>%
  gt() %>%
  tab_header(
    title = "Average Review Rating by Category"
  ) %>%
  fmt_number(
    columns = "Average_Review_Rating",
    decimals = 2
  ) %>%
  cols_label(
    Average_Review_Rating = md("**Average Rating**") ,
    Category = md("**Category**")
  )
```

### Level 4
Data visualization 
```{r lvl 4 viz, warning=FALSE}
ggplot(naRMData, aes(x = `Review Rating`, y = `Purchase Amount (USD)`)) +
  geom_point(width = 0.2, alpha = 0.7, colour = "#B9C0DA") +  
  geom_jitter(width = 0.2, alpha = 0.7, colour = "#B9C0DA") +  
  geom_smooth(method = "loess", colour = "#F8333C", se = TRUE, size = 1.7) +
  labs(
    title = "Relationship of Purchase Amount & Review Rating",
    x = "Review Rating",
    y = "Purchase Amount (USD)"
  ) +
  theme_minimal(base_size = 15) +  
  theme(
    panel.background = element_rect(fill = "black", colour = "black"),  
    plot.background = element_rect(fill = "black", colour = "black"),   
    panel.grid.major = element_line(colour = "white"),                
    panel.grid.minor = element_line(colour = "white"),                
    axis.text = element_text(colour = "#FFC53A"),                       
    axis.title = element_text(colour = "#FFC53A"),                      
    plot.title = element_text(colour = "#51E5FF", hjust = 0.5, size = 19))

# corr of vars 
cor_test <- cor.test(naRMData$`Purchase Amount (USD)`, naRMData$`Review Rating`, use = "complete.obs")
```
#### Correlation Results

>**Correlation Coefficient (r):** `r round(cor_test$estimate, 3)`

>**P-Value of Correlation:** `r round(cor_test$p.value, 3)`

---

---

---