---
title: "PsiChi R Coding Competition Submission--October"
author: "Brady M. Chisholm"
date: "October, 2024"
output: pdf_document
---

## Load Data & packages

### Assume user has nothing installed, support this in .rmd file

```{r}

# List required packages, install if needed
required_packages <- c("knitr","formatR","tidyverse", "ggplot2", "tidyr", "dplyr", 
                       "lubridate", "scales", "patchwork")

suppressMessages(suppressWarnings({
  for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE)) {
    install.packages(pkg, repos = "https://cloud.r-project.org")
  }
  library(pkg, character.only = TRUE)}  # end for 
})) # end suppress 
knitr::opts_chunk$set(comment = '', fig.width = 8, fig.height = 6, warning = FALSE, 
                  tidy.opts = list(width.cutoff = 60))

# use osf url. Tried to source from my github but repo errors 
horrorDataSrc <- "https://osf.io/download/mxs49/?direct&mode=render"
```

## Data Processing -- Level 1 

### Filter movies with missing values 

```{r level 1 pt 1}
# read in data from src 
HorrorIMDB <- read.csv(horrorDataSrc)

# Filter missing value in budget, runtime, parentalrating, and rating column 
cleanedMovies <- HorrorIMDB %>%
  filter(!is.na(Budget), !is.na(RunTime), !is.na(ParentalRating), !is.na(Rating))

# Separate the 'Genre' column into multiple genre categories
genreSorted <- cleanedMovies %>%
  separate_rows(Genres, sep = "\\|") %>%
  mutate(Genre = trimws(Genres)) # trim white space created 
```

## Descriptive Statistics -- Level 2

### Create summary statistic script 

```{r descriptive stats pt 1}
# Avg, std, median, range of rating variable 
summaryStat <- function(myDat){
  avg <- mean(myDat, na.rm = TRUE)
  std <- sd(myDat, na.rm = TRUE)
  med <- median(myDat, na.rm = TRUE)
  rg <- range(myDat, na.rm = TRUE)
  # organize results and make them print in a pretty way. Alt a data frame? 
  stats_df <- data.frame(
    Descriptive_Stat = c("Average", "Standard Deviation", "Median", "Range"),
    Value = c(round(avg, 2), round(std, 2), round(med, 2), 
              paste("(", rg[1], " to ", rg[2], ")", sep = ""))) # clean up for printing 
  
  print(stats_df, row.names = FALSE)
  
  # suppress output, just print cleaned df 
  invisible(list(avg = avg, std = std, med = med, range = rg))} # function end 

summaryStat(HorrorIMDB$Rating)
```

### Show average by genre

```{r descriptive stats pt 2}
# Define the function to calculate the average rating by main genre
genreAvg <- function(data) {
  # uses dyplr, ensure install earlier in script 
  
  # Check for required columns in data
  if (!all(c("Genre","Genre", "Rating") %in% colnames(data))) { #FIXME %in% ?? 
    # throw error if we don't have enough info 
    stop("Data must contain genre and rating data")}
  
  # Calculate rating for each genre 
  genre_avg_df <- data %>% # FIXME double pipe? 
    group_by(Genre) %>% # use genre as grouper 
    summarise(Average_Rating = mean(Rating, na.rm = TRUE)) %>%
    arrange(desc(Average_Rating)) # Sort by average rating, highest first 
  
  # Print pretty results
  print(genre_avg_df, row.names = FALSE)
  
  # return table but suppress 
  invisible(genre_avg_df)} # function end 
```

#### Test function 

```{r}
# test function 
genreAvg(genreSorted) # function only works when called on sorted data 
```

## Data Visualization -- Level 3 

### Plot the budget for movies over time

```{r data viz. data wrangle}
# clean up data first. Need to extract year for plotting
HorrorIMDB <- HorrorIMDB %>%
  mutate(Release_Date = mdy(Release_Date),  # switch to date
         Year = year(Release_Date),         # get year 
         Budget_Million = Budget / 1e6) #%>%  # convert budg to millions

# bad code. Filter was removing all low values, no data left to plot. Fixed  
HorrorIMDB <- HorrorIMDB %>%
  filter(!is.na(Year))

# split data
early_data <- HorrorIMDB %>% filter(Year <= 1905)
later_data <- HorrorIMDB %>% filter(Year >= 1910)
```

### After cleaning, finally can plot

```{r data visualization}
# Graph budget of movies  over time 
# first plot early data 
early_plot <- ggplot(early_data, aes(x= Year,y =  Budget_Million)) +
  geom_jitter(color = "black", alpha = 0.7, width = 0.25, size = 2) + # only one yr, distribute points
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", suffix = "M")) +
  labs(
    title = "Movie Budgets in 1905",
    x = "Year - 1905   ", 
    y = "Budget in Millions $") +
  theme_minimal() + 
  theme(
    text = element_text(family = "serif", color = "black",size = 12, hjust = .5, face = "bold"),
    panel.grid = element_line(color = "orange"),
    panel.grid.major.x = element_line(color = "orange"),
    panel.grid.major.y = element_line(color = "black"),
    axis.ticks.y = element_line(color = "black"),
    axis.ticks.x = element_line(color = "black"),
    plot.title = element_text(face = "bold", size = 9, hjust = .95, color = "black"),
    plot.background = element_rect(fill = "orange"),
    axis.text.y = element_text(family = "serif",color = "black",face = "bold"),
    #FIXME make x axis text disappear 
    axis.text.x = element_blank(),
    axis.line.x = element_line(color = "black"),
    axis.line.y = element_line(color = "black")
  ) # close theme 

later_plot <-ggplot(later_data, aes(x = Year, y = Budget_Million)) +
  geom_smooth(method = "glm", linewidth = 1.5, color = "#00CED1", fill = "orange", 
              na.rm = TRUE, se = TRUE,span = 1, level = .98) +
  geom_jitter(color = "white", alpha = 0.55, width = 0.47, size = 2) + 
  scale_y_continuous(labels = scales::dollar_format(prefix = "$", suffix = "M")) +
  labs(
    title = "Movie Budgets from 2012 to 2017",
    x = "Year (2012-2017)",
    y = NULL) +
  theme_minimal() +
  theme(
    axis.text = element_text(color = "white"),
    text = element_text(family = "serif", color = "black",size = 12, face = "bold"),
    plot.title = element_text(face = "bold", size = 12, hjust = 0.5, color = "white"),
    axis.text.x = element_text(angle = 45, hjust = 0.5, color = "white"),
    axis.title.x = element_text(face = "bold", hjust = 0.5, color = "white"),
    panel.grid = element_line(color = "#8B0000", linewidth = 0.1),
    plot.background = element_rect(fill = "black")) 

# combine plots, add final aesthetics 
comboPlot <- early_plot + later_plot + plot_layout(ncol = 2, widths = c(0.3, 2.5))
comboPlot + 
  plot_annotation(
    title = "Halloween Movie Budgets Over Time",
    subtitle = "Comparing Budgets from 1905 to Modern Day; 2012-2017",
    theme = theme(
      text = element_text(family = "serif", color = "ghostwhite"),
      plot.title = element_text(face = "bold", size = 18, color = 'maroon', hjust = 0.5),
      plot.subtitle = element_text(face = 'italic', size = 14, color = 'red', hjust = 0.5)))
```

## Inferential Statistics -- Level 4

### Do horror movies get better ratings than mystery movies? 

```{r inferential pt 1}
# Do movies with bigger budgets get better ratings?? 
cleaned_data <- HorrorIMDB %>%
  filter(!is.na(Budget), !is.na(Rating))

ggplot(cleaned_data, aes(x = Budget, y = Rating)) +
  geom_point(alpha = 0.7, color = "#32CD32", size = 2) +  # Neon green points
  geom_smooth(method = "lm", color = "#FF7518", se = TRUE, size = 1.1) +  # Orange regression line
  scale_x_continuous(labels = scales::dollar_format(prefix = "$", suffix = "M")) +
  labs(
    title = "Movie Budget and Rating: Halloween Edition",
    x = "Budget ($ Million)",
    y = "IMDB Rating") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "orange"),
    axis.title = element_text(size = 14, color = "white"),
    axis.text = element_text(size = 12, color = "orange"),
    axis.line.x = element_line(color = "white"),
    axis.line.y = element_line(color = "white"),
    panel.background = element_rect(fill = "#222222", color = NA),  # Dark panel background
    plot.background = element_rect(fill = "#222222", color = NA),   # Dark plot background
    plot.margin = margin(15, 15, 15, 15)  # Adds space around plot for readability
  )
# find our correlation statistic  
correlation <- cor(cleaned_data$Budget, cleaned_data$Rating, use = "complete.obs")
print(paste("Correlation between budget & rating is", round(correlation, 2)))

# simple linear model
model <- lm(Rating ~ Budget, data = cleaned_data)
summary(model)
```

The correlation between budget and rating is **0.13**. This indicates a **weak positive correlation** between budget and IMDB rating. Results suggest that higher budget may be weakly associated with better ratings, but the effect is not strong. 

A linear regression model was fitted to explore the effect of budget on IMDB ratings. The model summary is:

- **Intercept**: 4.97
- **Budget Coefficient**: 2.13e-08
- **t-value**: 3.72
- **p-value**: 0.000217 (**p < 0.001**)

The regression equation used is:

\[
\text{Rating} = 4.97 + (2.13 \times 10^{-8}) \times \text{Budget}
\]

The p-value of 0.000217 tells us a statistically significant relationship between budget and ratings exists. However, the **R-squared value** is **0.016**. A low R-squared value such as this says budget explains only about 1.6% of the variance in movie ratings. This tells us that there are possibly many other factors influencing movie success as measured by ratings.

### Do horror movies get better ratings than mystery? 

```{r inferential pt 2}
print("Genres in dataset:")
print(unique(genreSorted$Genres))

# filter out genres in question 
cleanGenreDat <- genreSorted %>%
  filter(!is.na(Rating), Genres %in% c(" Horror", " Mystery")) # leading space?!?

# assert data existence
unique_genres <- unique(cleanGenreDat$Genres)
print("Genres present in the filtered dataset:")
print(unique_genres)
if (length(unique_genres) < 2) {
  stop("Error: Dataset must contain both 'Horror' and 'Mystery' genres for comparison.")}

# plot the ratings 
ggplot(cleanGenreDat, aes(x = Genres, y = Rating, fill = Genres)) +
  geom_boxplot(
    alpha = 0.8, 
    outlier.color = 'limegreen', 
    outlier.shape = 21, 
    outlier.size = 3, 
    color = "orange",  
    size = 1.0
  ) +
  scale_fill_manual(values = c(" Horror" = "#FF7518", " Mystery" = "#6A0DAD")) +  # Halloween colors
  labs(
    title = "Horror vs Mystery Movies",
    subtitle = "Spooky Themed",
    x = "Genre",
    y = "IMDB Rating"
  ) +
  theme_minimal() +
  theme(
    #FIXME plot axis as white to match grid
    plot.title = element_text(size = 18, face = "bold", color = "orange",hjust = 0.5),
    axis.title = element_text(size = 14, color = "white"),
    axis.text = element_text(size = 12, color = "orange"),
    plot.subtitle = element_text(size = 13, face = "italic", color = "orange",hjust = 0.5),
    legend.position = "none",
    plot.background = element_rect(fill = "#222222", color = NA),  # Dark background
    panel.background = element_rect(fill = "#222222"),
    text = element_text(family = "serif"))  # end theme 

# calculate mean stat
genre_means <- cleanGenreDat %>%
  group_by(Genres) %>%
  summarise(Mean_Rating = mean(Rating, na.rm = TRUE))

print("Mean Ratings for Each Genre:")
print(genre_means)
```
The average rating for each genre is:
- **Horror**: 6.3
- **Mystery**: 7.1

Analysis suggests that **Mystery movies tend to receive higher ratings on average** than **horror** movies, and this difference is statistically significant.

```{r,include = FALSE,echo = TRUE}
invisible(cat(" /\\_/\\\n( o.o )\n > ^ <\n"))
```
