---
title: "januaryR-Code"
output: html_document
author: "Brady M. Chisholm"
---
### Prep & Setup
```{r setup, echo=TRUE, message=FALSE, warning=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org"))
# install.packages("knitr", quiet = TRUE)
library(knitr)
library(dplyr)
required_packages <- c("ggplot2","tidyr","readr","moments","psych","gt","dplyr") 
for (pkg in required_packages) {
  if (!require(pkg, character.only = TRUE)) { 
    install.packages(pkg)               
    library(pkg, character.only = TRUE)} 
  else {
    library(pkg, character.only = TRUE)}}
```
Load data 
```{r data loading, echo=TRUE, message=FALSE, warning=FALSE}
# Load CSV file
datLink <- "https://raw.githubusercontent.com/bchisholm08/OSF_graphingCompetition/refs/heads/main/Month/January_25/datasets/world_cheese_awards_2021.csv"

cheeseData <- read_csv(datLink)
# preview data / var types
head(cheeseData)
colnames(cheeseData)
```

### Level 1 
remove any rows missing a value for company, product name, and/or rating

check if we have NaN or NA vals 
```{r}
naRMData <- cheeseData %>%
  filter(if_all(c("company", "product_name", "rating"), ~ !is.na(.)))

cat("Original Dimensions: ", dim(cheeseData), "\n")
cat("Filtered Dimensions: ", dim(naRMData), "\n")

# make sure we have no NaNs left
sapply(naRMData[c("company", "product_name", "rating")], function(x) sum(is.na(x)))

```

```{r check vals}
dim(naRMData)
dim(cheeseData)
```

### Level 2 
Make a table to show the count of each category within rating, and then make another table showing counts for each country 
```{r}
catINRating %>%
  gt() %>%
  tab_header(
    title = "Count of Categories by Rating"
  ) %>%
  fmt_number(
    columns = 
    decimals = 2
  ) %>% 
  cols_label(
    
  )
  
```

descriptiveStats %>%
  gt() %>%
  tab_header(
    title = "Descriptive Statistics for Review Rating"
  ) %>%
  fmt_number(
    columns = "Value",
    decimals = 3
  ) %>%
  cols_label(
    Metric = md("**Statistic**"),
    Value = md("**Value**"))


### Level 3 
Make a table showing how many of each (rating) each (Category_Corrected). Which cheese has the most “Super Golds?” Next, modify the table and indicate which country produced the most “Super Gold” cheeses
```{r}

```

### Level 4 
Generate a map that shows which countries have won super gold medals
```{r}

```

